Auditory Scene Similarity Retrieval and Classification with Relevance-based Quantization of Scattering Features

### 2016-12-16 paper/structureScene.tex
New intro
This is currently too long, I think, so will need to be condensed
to relevant size.

\IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has grown considerably over the past decades.
In order to measure the effect of human activity and climate change on animal biodiversity, researchers in the field of ecoacoustics \cite{krause} have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}. 
In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
Because they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for the signal processing community.

Ecoacoustics is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science. Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
As a consequence, few well-designed datasets are readily available for evaluation purposes.

A closely related, and more mature, field of investigation is the classification of acoustic scenes, where the aim is to predict labels by processing audio data.
While narrower in its range of applications, it has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of well-designed datasets.

Despite the differences, ecoacoustics and acoustic scene classification (ASC) task both raise the following question: how can we efficiently and effectively process a large amount of audio data using representations that are compact, generic in their design, and flexible in their use?
The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts for environmental audio processing: 1) application of the wavelet scattering transform to extract discriminative features and 2) large time-scale modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on a few salient sound objects that carry most of the semantic information.

>>>>

\IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has grown considerably over the past decades.
In order to measure the effect of human activity and climate change on animal biodiversity researchers have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}. 
In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
Because they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for the signal processing community.

This works falls in the realm of ecoacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}. Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
Consequently, few well-designed datasets are readily available for evaluation purposes.

An important problem in this field is the characterization of auditory scenes for the purposes of similarity estimation. One application of this is acoustic scene similarity retrieval (ASSR), where the scenes judged ``most similar'' to a given query scene are extracted from some larger dataset. This requires the auditory scenes to be represented in some way that captures their distinctive qualities. Often, this has been achieved by bag-of-frames approaches, which attempt to describe the auditory scene recording using summary statistics. Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain rare, distinct sound events that discriminate the scene from others.Indeed, experiments in cognitive psychology suggests that this matches more closely the way that humans and other animals perceive acoustic scenes. \ja{Refs?}

To design better representations for the similarity retrieval task, we draw inspiration from a closely related, and more mature, field of investigation: the classification of acoustic scenes.This task, consisting of predicting scene labels from audio recordings, is narrower in its range of applications, but has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive,guastavino_ideal_2006}. In addition, it has received significant attention from the data mining community for some time, resulting in the availability of well-designed datasets and competitive algorithms for solving it.

The first question to be answered when confronted with this task is how is the signal to be represented at a low level. In other words, how do we represent an extract of the auditory scene signal over a sub-second interval? A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years \cite{logan}. \ja{Other MFCC works?} A richer set of features are found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.

The next question is the choice of a classifier. Here, a class of particularly successful algorithms train a classifier, such as a support vector machine (SVM), to discriminate between short extracts of an auditory scene, and at the classification stage divide the signal into short extracts which are classified separately before pooling the classification results to reach a final decision. This last step is known as late integration and is crucial to the success of these methods, since it allows the classifier to concentrate on the short, distinct sound events that discriminate between scenes. The success of the late integration technique again leads credence to the idea that rare, distinct events in an auditory scene are sufficient to characterize it completely. Combining scattering transforms for low-level representation with late-integrated SVMs used for classifying the scene at a higher level, we obtain a competitive auditory scene classification system, demonstrating the usefulness of both concepts for this task.

Returning to scene retrieval setting, we cannot use supervised models such as SVMs, but we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization. We therefore propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a number of clusters. We then use these to define a set of distances that can subsequently be used for similarity retrieval.Testing this approach on an scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models, applied both to MFCCs and scattering coefficients.
